Clines from PDE

### 2015-10-23 notes/pde-clines.Rmd
.

We know that 
$$\begin{aligned}
q(x,A,0,r) = 1  \\
q(x,B,0,r) = 0  \\
\end{aligned}$$
and that $\partial_t q = G q$, i.e.,
(writing $q(x,A,t,r) = q_A(x,t)$ for the moment):
$$\begin{aligned}
    \partial_t q_A(x,t)
        &= 
        \grad \log(p(x)) \cdot \grad q_A(x,t) + \frac{\sigma^2}{2} \Delta q_A(x,t) + r (1-p(x)) (q_B(x,t)-q_A(x,t)) \\
    \partial_t q_B(x,t)
        &= 
        \grad \log(1-p(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r p(x) (q_A(x,t)-q_B(x,t)) \\
\end{aligned}$$

Let's solve this numerically.
```{r solve_pde_setup, cache=TRUE}
s <- 0.1
tau <- 100
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
cline.A <- setup.prop.1D(grid=thegrid, func=tanh_cline, s=s)
cline.B <- setup.prop.1D(grid=thegrid, func=function(x,s)1-tanh_cline(x,s), s=s)
yinit <- c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) )
tt <- seq(0,tau,length.out=100)
r <- 1
pde.fn <- function (t,y,parms,...) {
    yA <- y[1:thegrid$N]
    yB <- y[thegrid$N+(1:thegrid$N)]
    tran.A <- tran.1D(C=yA, A=cline.A, D=1/2, dx=thegrid)$dC
    tran.B <- tran.1D(C=yB, A=cline.B, D=1/2, dx=thegrid)$dC
    list( c( 
            tran.A + r * cline.B$mid * (yB-yA),
            tran.B + r * cline.A$mid * (yA-yB)
            ) )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends="solve_pde_setup"}
A.inds <- 1+(1:thegrid$N)
B.inds <- 1+thegrid$N+(1:thegrid$N)
pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*cline.A$mid[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*cline.B$mid[col(pde.soln[,B.inds])] )
attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
attr(pde.cline,"nspec")<-1
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
# same thing, plotted with lines
clineplot(pde.cline,main=sprintf("clines, at r=%0.2f",r))
```

>>>>

We know that 
$$\begin{aligned}
q(x,A,0,r) = 1  \\
q(x,B,0,r) = 0  \\
\end{aligned}$$
and that $\partial_t q = G q$, i.e.,
(writing $q(x,A,t,r) = q_A(x,t)$ for the moment):
$$\begin{aligned}
    \partial_t q_A(x,t)
        &= 
        \grad \log(p(x)) \cdot \grad q_A(x,t) + \frac{\sigma^2}{2} \Delta q_A(x,t) + r (1-p(x)) (q_B(x,t)-q_A(x,t)) \\
    \partial_t q_B(x,t)
        &= 
        \grad \log(1-p(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r p(x) (q_A(x,t)-q_B(x,t)) .
\end{aligned}$$

Paving the way for analytic approximations
------------------------------------------

Note that $p$ appears twice here:
first pushing lineages, and second causing recombination.
We may want to decouple these for the purposes of approximation:
we'll consider the following system of PDE:
$$\begin{aligned}
    \partial_t q_A(x,t)
        &= 
        \grad \log(u(x)) \cdot \grad q_A(x,t) + \frac{\sigma^2}{2} \Delta q_A(x,t) + r (1-v(x)) (q_B(x,t)-q_A(x,t)) \\
    \partial_t q_B(x,t)
        &= 
        \grad \log(1-u(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r v(x) (q_A(x,t)-q_B(x,t)) .
\end{aligned}$$

Here's a function to solve this numerically:
```{r solve_pde_setup, cache=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
solve.pde <- function ( u, v=u, r, times, ... ) {
    vfn <- function(x,t){v(x=x,t=t,...)}
    cline.A <- setup.prop.1D(grid=thegrid, func=u, ...)
    cline.B <- setup.prop.1D(grid=thegrid, func=function(x,...)1-u(x,...), ...)
    yinit <- c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) )
    r <- 0.1
    pde.fn <- function (t,y,parms,...) {
        yA <- y[1:thegrid$N]
        yB <- y[thegrid$N+(1:thegrid$N)]
        tran.A <- tran.1D(C=yA, A=cline.A, D=1/2, dx=thegrid)$dC
        tran.B <- tran.1D(C=yB, A=cline.B, D=1/2, dx=thegrid)$dC
        list( c( 
                tran.A + r * (1-vfn(x=thegrid$x.mid,t=t)) * (yB-yA),
                tran.B + r * vfn(x=thegrid$x.mid,t=t) * (yA-yB)
                ) )
    }
    ode.1D( y=yinit, times=times, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
}
```

Numerical solutions
-------------------

Let's solve this numerically.
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
s <- 0.1
tau <- 100
tt <- seq(0,tau,length.out=100)
r <- 0.1
pde.soln <- solve.pde( u=tanh_cline, r=r, times=tt, s=s )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends="solve_pde_setup"}
A.inds <- 1+(1:thegrid$N)
B.inds <- 1+thegrid$N+(1:thegrid$N)
pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*cline.A$mid[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*cline.B$mid[col(pde.soln[,B.inds])] )
attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
attr(pde.cline,"nspec")<-1
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
# same thing, plotted with lines
clineplot(pde.cline,main=sprintf("clines, at r=%0.2f",r))
```
